---
title: "NK_Scoring"
author: "Robin Wagner (UoM Laptop)"
date: "2023-08-23"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Introduction

In this code I will load and process the raw counts data generated from CRLM tissue and perform single-sample scoring against a CRC-specific NK signature using singscore. 

### Setup 

```{r}
 
library(dplyr) 
library(edgeR) 
library(singscore)
library(ggplot2)
library(RColorBrewer)

```


### Prepare data

Read in and prepare the raw counts data for the 58 samples. 

```{r}

expr <- readRDS("data/Hollande_DGEList_Tum_Raw.RDS"); dim(expr)

```

Remove duplicated sample. 

```{r} 

expr$samples <- expr$samples[-12,]
expr$counts <- expr$counts[,-12]
dim(expr$samples); dim(expr$counts)
all(colnames(expr$counts) == rownames(expr$samples))

```

Filter out primary tumours as these are irrelevant to our analysis. 

```{r}

expr$samples <- expr$samples %>% filter(!Tissue == "CoT"); summary (expr$samples$Tissue)
expr$counts <- expr$counts[,rownames(expr$samples)]; dim(expr$counts)

```

Load the NK signature now so that we preserve these genes upon filtering. 

```{r}

nk_sig <- read.csv("data/20230822_Shembrey_NK_Signature.csv"); head(nk_sig)

```

Filter out lowly-expressed genes. 

```{r}

keep <- filterByExpr(expr$counts)
keep2 <- rownames(expr$counts) %in% nk_sig$ENTREZID
expr_filt <- expr[keep | keep2, ]; dim(expr_filt)

```

Normalise for library size. 

```{r}

expr_norm <- calcNormFactors(expr_filt, method = "TMM")

```

### Perform NK scoring

Normalise read counts for gene length, in order to perform sample scoring with singscore. 

```{r}

expr_rpkm <- edgeR::rpkm(expr_norm, gene.length = expr_norm$genes$Length, log = T)

```

Rank genes. 

```{r}

expr_ranked <- rankGenes(expr_rpkm)

```

Score samples using NK signature. 

```{r}

score <- simpleScore(expr_ranked, nk_sig$ENTREZID)[,1]; head(score)

nk_score <- data.frame(sample = colnames(expr_rpkm),
                       score = score, 
                       quartile = ntile(score, 4)) %>%
  mutate(designation = ifelse(quartile == 1, "low",
                              ifelse(quartile == 4, "high", "intermediate"))); head(nk_score)

```

Visualise the scores. 

```{r, fig.dim = c(6,20)}

ggplot(nk_score, aes(x = reorder(sample, score), y = score, fill = designation)) +
  geom_bar(stat = "identity") +
  coord_flip() + 
  theme_bw() + 
  scale_fill_manual(values = brewer.pal(3, "Paired")) + 
  labs(x = "") +
  theme(text = element_text(size = 15))

```
